---
title: "Stats525HW7"
author: "Luke Geel"
date: "3/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2.4
a. Obtain a 99 percent confidence interval for b1. Interpret your confidence interval. Does it
include rero? Why might the director of admissions be interested in whether the confidence
interval includes zero?
CI:(0.00175,0.094787) This means that we can be 95% confident that for every increase in GPA, the ACT score will increase by between 0.00175 and 0.094787. This does not include zero. The director would be interseted in whether the confidence interval includes zero because if it did then there might not be a correlatin between ACT scores and GPA.
```{r}
library(tidyverse)
loadRData <- function(fileName){
  load(fileName)
  get(ls()[ls() != "fileName"])
}
gpa <- loadRData("/Users/lukegeel/Downloads/gpa_spring2021.RData")
n <- nrow(gpa) 
y <- gpa$Y 
x <- gpa$X 
linmod <- lm(y~x) 
b1 <- linmod$coef[2]
s.b1 <- summary(linmod)$coef[2, 2]
alpha <- 0.01
qt <- qt(alpha/2, n - 2)
pvalue <- pt(-abs(b1/s.b1),n-2)+(1-pt(abs(b1/s.b1),n-2))
lower <- b1+s.b1*qt(alpha/2,n-2)
upper <- b1-s.b1*qt(alpha/2,n-2)
lower
upper
```
b. Test, using the test statistic t*, whether or not a linear association exists between student's
ACT score (X) and GPA at the end of the freshman year (Y). Use a level of significance of
.Ol. State the alternatives, decision rule, and conclusion.
Alternative hypothesis: There is no linear association between student's ACT score and GPA at the end of the freshman year.
Decision rule: If the test statistic exceeds the t-value given the alpha level, we reject the alternative hypothesis.
Conclusion: The test statistci exceeds t, so we can reject the alternative hypothesis and conclude that there is a linear association between ACT scores and GPA.
```{r}
library(tidyverse)
loadRData <- function(fileName){
  load(fileName)
  get(ls()[ls() != "fileName"])
}
gpa <- loadRData("/Users/lukegeel/Downloads/gpa_spring2021.RData")
n <- nrow(gpa) 
y <- gpa$Y 
x <- gpa$X 
linmod <- lm(y~x) 
b1 <- linmod$coef[2]
s.b1 <- summary(linmod)$coef[2, 2]
alpha <- 0.01
tquantile <- qt(1 - alpha/2, n - 2)
statistic <- b1/s.b1
tquantile
statistic
```
c. What is the P-value of your test in part (b)? How does it support the conclusion reached in
part (b)? 
P-value: 0.007   The P-value is less than the alpha level so it supports our conclusion that a linear association exists.
```{r}
library(tidyverse)
loadRData <- function(fileName){
  load(fileName)
  get(ls()[ls() != "fileName"])
}
gpa <- loadRData("/Users/lukegeel/Downloads/gpa_spring2021.RData")
n <- nrow(gpa) 
y <- gpa$Y 
x <- gpa$X 
linmod <- lm(y~x) 
b1 <- linmod$coef[2]
s.b1 <- summary(linmod)$coef[2, 2]
alpha <- 0.01
tquantile <- qt(1 - alpha/2, n - 2)
pvalue <- 2*pt(-abs(b1/s.b1), n - 2)
pvalue

```

2.5. Refer to Copier maintenance Problem 1.20.
a. Estimate the change in the mean service time when the number of copiers serviced increases
by one. Use a 90 percent confidence interval. Interpret your confidence interval.
We can be 90% confident that when the number of copiers serviced increases by one, the service time increases by between 14.0258 and 16.08582 minutes.
```{r}
data <- loadRData("/Users/lukegeel/Downloads/copier_spring2021.RData")
n <- nrow(data) 
y <- data$Y 
x <- data$X 
linmod <- lm(y~x) 
b1 <- linmod$coef[2]
s.b1 <- summary(linmod)$coef[2, 2]
alpha <- 0.1
qt <- qt(alpha/2, n - 2)
pvalue <- pt(-abs(b1/s.b1),n-2)+(1-pt(abs(b1/s.b1),n-2))
lower <- b1+s.b1*qt(alpha/2,n-2)
upper <- b1-s.b1*qt(alpha/2,n-2)
lower
upper
```
b. Conduct a t test to determine whether or not there is a linear association between X and Y
here; control the alpha risk at .10. State the alternatives, decision rule, and conclusion. What is
the P-value of your test?
Alternative hypothesis: There is not a linear association between X and Y.
Decision rule: If the test statistic exceeds the t-value given the alpha level, we reject the alternative hypothesis.
Conclusion: The test statistic doesn't exceed t, so we fail to reject the alternative hypothesis and say that there isn't a linear association between X and Y.
P-value: 6.093652e-27 
```{r}
data <- loadRData("/Users/lukegeel/Downloads/copier_spring2021.RData")
n <- nrow(data) 
y <- data$Y 
x <- data$X 
linmod <- lm(y~x) 
b1 <- linmod$coef[2]
s.b1 <- summary(linmod)$coef[2, 2]
alpha <- 0.1
tquantile <- qt(1 - alpha/2, n - 2)
statistic <- b1/s.b1
pvalue <- 2*pt(-abs(b1/s.b1), n - 2)
pvalue
tquantile
statistic
```
c. Are your results in parts (a) and (b) consistent? Explain.
No, the results aren't consistent. From a, we get there the mean change is about 15 minutes for each additional copier but from b we concluded that there isn;t a linear relationship. These results contradict each other.
```{r}
```
e. Does bo give any relevant information here about the "start-up" time on calls-Le., about
the time required before service work is begun on the copiers at a customer location?
No, b0 does not give any relevant information about the start-up time on calls.
```{r}

```

1.44. Refer to the CDI data set in Appendix C.2.
a. For each geographic region, regress per capita income in a CDr (Y) against the percentage of individuals in a county having at least a bachelor's degree (X). Assume that first-order regression model (1.1) is appropriate for each region. State the estimated regression functions.
Region 1: 9223.8 + 522.2X
Region 2: 13581.4 + 238.7X
Region 3: 10529.8 + 330.6
Region 4: 8615.1 + 440.3X
```{r}
cdi <- loadRData("/Users/lukegeel/Downloads/cdi_spring2021.RData")
region1 <- cdi%>%filter(X17 == 1)
n <- nrow(region1)
Y <- region1$X15 
X <- region1$X12 
linmod1 <- lm(Y~X)
linmod1
region2 <- cdi%>%filter(X17 == 2)
n <- nrow(region2)
Y <- region2$X15 
X <- region2$X12 
linmod2 <- lm(Y~X)
linmod2
region3 <- cdi%>%filter(X17 == 3)
n <- nrow(region3)
Y <- region3$X15 
X <- region3$X12 
linmod3 <- lm(Y~X)
linmod3
region4 <- cdi%>%filter(X17 == 4)
n <- nrow(region4)
Y <- region4$X15 
X <- region4$X12 
linmod4 <- lm(Y~X)
linmod4

```
b. Are the estimated regression functions similar for the four regions? Discuss.
Yes, the estimated regression functions are relatively similar for all four regions. The intercepts are all between 8000 and 14000 and the slope of X are all between 230 and 520.
```{r}

```
c. Calculate MSE for each region. Is the variability· around the fitted regression line approximately the same for the four regions? Discuss. 
Region 1 MSE: 7335008
Region 2 MSE: 4411341
Region 3 MSE: 7474349
Region 4 MSE: 8214318
I believe that the variability around the fitted regression line is approximately the same for the four regions as all the MSEs are very large, all between 4 and 8 million.
```{r}
loadRData("/Users/lukegeel/Downloads/cdi_spring2021.RData")
s1 <- summary(linmod1)$s
s1^2
s2 <- summary(linmod2)$s
s2^2
s3 <- summary(linmod3)$s
s3^2
s4 <- summary(linmod4)$s
s4^2

```

2.63 Refer to the CDI data set in Appendix C.2 and Project l.44. Obtain a separate interval estimate
of b1, for each region. Use a 90 percent confidence coefficient in each case. Do the regression
lines for the different regions appear to have similar slopes?
Region 1 b1 CI: (460.5177,583.8)
Region 2 b1 CI: (193.4858,283.853)
Region 3 b1 CI: (285.7076,375.5158)
Region 4 b1 CI: (364.7585,515.8729)
Yes, the regression lines for the different regions appear to have similar slopes.
```{r}
n <- nrow(region1)
Y <- region1$X15 
X <- region1$X12 
linmod1 <- lm(Y~X)
b1 <- linmod1$coef[2]
s.b1 <- summary(linmod1)$coef[2, 2]
alpha <- 0.1
qt <- qt(alpha/2, n - 2)
pvalue <- pt(-abs(b1/s.b1),n-2)+(1-pt(abs(b1/s.b1),n-2))
lower <- b1+s.b1*qt(alpha/2,n-2)
upper <- b1-s.b1*qt(alpha/2,n-2)
lower
upper
```
```{r}
n <- nrow(region2)
Y <- region2$X15 
X <- region2$X12 
linmod2 <- lm(Y~X)
b1 <- linmod2$coef[2]
s.b1 <- summary(linmod2)$coef[2, 2]
alpha <- 0.1
qt <- qt(alpha/2, n - 2)
pvalue <- pt(-abs(b1/s.b1),n-2)+(1-pt(abs(b1/s.b1),n-2))
lower <- b1+s.b1*qt(alpha/2,n-2)
upper <- b1-s.b1*qt(alpha/2,n-2)
lower
upper
```
```{r}
n <- nrow(region3)
Y <- region3$X15 
X <- region3$X12 
linmod3 <- lm(Y~X)
b1 <- linmod3$coef[2]
s.b1 <- summary(linmod3)$coef[2, 2]
alpha <- 0.1
qt <- qt(alpha/2, n - 2)
pvalue <- pt(-abs(b1/s.b1),n-2)+(1-pt(abs(b1/s.b1),n-2))
lower <- b1+s.b1*qt(alpha/2,n-2)
upper <- b1-s.b1*qt(alpha/2,n-2)
lower
upper
```
```{r}
n <- nrow(region4)
Y <- region4$X15 
X <- region4$X12 
linmod4 <- lm(Y~X)
b1 <- linmod4$coef[2]
s.b1 <- summary(linmod4)$coef[2, 2]
alpha <- 0.1
qt <- qt(alpha/2, n - 2)
pvalue <- pt(-abs(b1/s.b1),n-2)+(1-pt(abs(b1/s.b1),n-2))
lower <- b1+s.b1*qt(alpha/2,n-2)
upper <- b1-s.b1*qt(alpha/2,n-2)
lower
upper
```

6.29. Refer to the CDI data set in Appendix C.2.
a. For each geographic region, regress the number of serious crimes in a CDI (Y) against
population density (X" total population divided by land area), per capita personal income
(X2 ), and percent high school graduates (X3 ). Use first-order regression model (6.5) with
three predictor variables. State the estimated regression functions.
Region 1: Y = -64466.231 + 17.383X1 - 1.406X2 + 1182.577X3
Region 2: Y = -4163.2673 + 33.6193X1 + 0.1024X2 - 2.7616X3
Region 3: Y = 38862.667 + 5.537 + 1.957X2 - 670.884X3
Region 4: Y = 129323.415 + 5.717X1 + 4.342X2 - 2159.920X3
```{r}
region1 <- cdi%>%filter(X17 == 1)
n <- nrow(region1)
Y <- region1$X10 
X1 <- region1$X5/region1$X4
X2 <- region1$X15 
X3 <- region1$X11 
linmod1 <- lm(Y~X1+X2+X3)
linmod1
region2 <- cdi%>%filter(X17 == 2)
n <- nrow(region2)
Y <- region2$X10 
X1 <- region2$X5/region2$X4
X2 <- region2$X15 
X3 <- region2$X11 
linmod2 <- lm(Y~X1+X2+X3)
linmod2
region3 <- cdi%>%filter(X17 == 3)
n <- nrow(region3)
Y <- region3$X10 
X1 <- region3$X5/region3$X4
X2 <- region3$X15 
X3 <- region3$X11 
linmod3 <- lm(Y~X1+X2+X3)
linmod3
region4 <- cdi%>%filter(X17 == 4)
n <- nrow(region4)
Y <- region4$X10 
X1 <- region4$X5/region4$X4
X2 <- region4$X15 
X3 <- region4$X11 
linmod4 <- lm(Y~X1+X2+X3)
linmod4
```
b. Are the estimated regression functions similar for the four regions? Discuss.
No, the estimated regression functions are not similar for the four regions. The intercepts are a wide range, some positive some negative. Also, the coefficients aren't always all positive and negative, they are different for each region.
```{r}
```
c. Calculate MSE and R2 for each region. Are these measures similar for the four regions?
Discuss.
Region 1: MSE: 787397480 R2: 0.8352049
Region 2: MSE: 1087623839 R2: 0.5285125
Region 3: MSE: 1367108167 R2: 0.09250839
Region 4: MSE: 6694591439 R2: 0.08665358
These measurements are not all similar for the four regions. The R2 values range form 0.83 to 0.08 so some are good fits some aren't.
```{r}
region1 <- cdi%>%filter(X17 == 1)
n <- nrow(region1)
Y <- region1$X10 
X1 <- region1$X5/region1$X4
X2 <- region1$X15 
X3 <- region1$X11 
linmod1 <- lm(Y~X1+X2+X3)
#linmod1
region2 <- cdi%>%filter(X17 == 2)
n <- nrow(region2)
Y <- region2$X10 
X1 <- region2$X5/region2$X4
X2 <- region2$X15 
X3 <- region2$X11 
linmod2 <- lm(Y~X1+X2+X3)
#linmod2
region3 <- cdi%>%filter(X17 == 3)
n <- nrow(region3)
Y <- region3$X10 
X1 <- region3$X5/region3$X4
X2 <- region3$X15 
X3 <- region3$X11 
linmod3 <- lm(Y~X1+X2+X3)
#linmod3
region4 <- cdi%>%filter(X17 == 4)
n <- nrow(region4)
Y <- region4$X10 
X1 <- region4$X5/region4$X4
X2 <- region4$X15 
X3 <- region4$X11 
linmod4 <- lm(Y~X1+X2+X3)
#linmod4
s1 <- summary(linmod1)$s
r.sq1 <- summary(linmod1)$r.sq
s1^2
r.sq1
s2 <- summary(linmod2)$s
r.sq2 <- summary(linmod2)$r.sq
s2^2
r.sq2
s3 <- summary(linmod3)$s
r.sq3 <- summary(linmod3)$r.sq
s3^2
r.sq3
s4 <- summary(linmod4)$s
r.sq4 <- summary(linmod4)$r.sq
r.sq4
s4^2
```

628. Refer to the CDI data set in Appendix C.2. You have been asked to evaluate two alternative
models for predicting the number of acti ve physicians (Y) in a CD!. Proposed model I includes
as predictor variables total population (X I), land area (X2 ), and total personal income (X}).
Proposed model II includes as predictor vcu-iables population density (X I, total population
divided by land area), percent of population greatef'lhan 64 years old (X2 ), and total personal
income (X}).
(a)Obtain a point estimate and 90% confidence interval for β1 under a normal errors linear regression
model that uses the number of active physicians in a CDI (Y ) as a response and total population (X1)
as a predictor.
CI: (0.002762953, 0.002928969)
```{r}
n <- nrow(cdi)
Y <- cdi$X8 
X <- cdi$X5 
linmod <- lm(Y~X)
b1 <- linmod$coef[2]
s.b1 <- summary(linmod)$coef[2, 2]
alpha <- 0.1
qt <- qt(alpha/2, n - 2)
pvalue <- pt(-abs(b1/s.b1),n-2)+(1-pt(abs(b1/s.b1),n-2))
lower <- b1+s.b1*qt(alpha/2,n-2)
upper <- b1-s.b1*qt(alpha/2,n-2)
lower
upper
```
(b) Obtain a point estimate and 90% confidence interval for β1 under a normal errors linear regression
model that uses the number of active physicians in a CDI (Y ) as a response and total population (X1),
land area (X2), and total personal income (X3) as predictors.
CI: (0.0004919585, 0.001485982)
```{r}
n <- nrow(cdi)
Y <- cdi$X8 
X1 <- cdi$X5 
X2 <- cdi$X4
X3 <- cdi$X16
linmod <- lm(Y~X1+X2+X3)
b1 <- linmod$coef[2]
s.b1 <- summary(linmod)$coef[2, 2]
alpha <- 0.1
qt <- qt(alpha/2, n - 2)
pvalue <- pt(-abs(b1/s.b1),n-2)+(1-pt(abs(b1/s.b1),n-2))
lower <- b1+s.b1*qt(alpha/2,n-2)
upper <- b1-s.b1*qt(alpha/2,n-2)
lower
upper
```
(c) Compare the interpretations of b1 in (a) and (c). Do they differ? If so, how?
In part a, its predicting the number of physicians based soley on total population where as in part b they number of physicians is predicted by total population, land area, and total personal income. 

(d) Compare the point estimates and 90% confidence intervals for β1 obtained in (a) and (b). Are they
identical?
They are not identical. The CI for part a is much smaller than part b. That means that when predicting the number of physicians using just the total population will produce more accurate results than using total population, land area, and total personal income.

(e) Explain how what you observe in (d) makes sense given what you concluded in (c).
Because we used more predictors in part b, the range of the CI would be larger as there are more variables in the prediction.






